### 2013-04-30 inst/examples/gaussian-process-basics.Rmd
include bayesian mcmc example

`ro comment=NA, message=FALSE, cache.path="gpb/", tidy=FALSE or`

```{r include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
```

@@

```{r results="hide"}
require(MASS)
require(reshape2)
require(ggplot2)
```

@@

>>>>

```{r plotting-options}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
library(ggplot2) # plotting 
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
```

@@

```{r results="hide"}
require(MASS)
require(reshape2)
```

@@

```{r}
lpriors <- function(pars){
   d.p <- c(5, 5)
  s2.p <- c(5, 5)

  prior <- unname(
    dgamma(exp(pars[1]), d.p[1], scale = d.p[2]) *
    dgamma(exp(pars[2]), s2.p[1], s2.p[2]) 
#    dunif(exp(pars[1]), 0, 100) * 
#    dunif(exp(pars[2]), 0, 100)
  )

  log(prior)
}

posterior <- function(pars, x, y){

  l <- exp(pars[1])
  sigma.n <- exp(pars[2])

  cov <- function(X, Y) outer(X, Y, SE, l)
  I <- diag(1, length(x))
  K <- cov(x, x)

  loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
    log(det(K + sigma.n^2*I)) -
    length(y) * log(2 * pi) / 2

  loglik + lpriors(pars)
}

```

```{r}
posterior(log(pars), obs$x, obs$y)
posterior(log(o$par), obs$x, obs$y)
```

```{r}
require(mcmc)
n <- 1e4
out <- metrop(posterior, log(pars), n, x = obs$x, y = obs$y)
out$accept
```

```{r}
postdist <- cbind(index=1:n, as.data.frame(exp(out$batch)))
names(postdist) <- c("index", names(pars))
df <- melt(postdist, id="index")
# TRACES
ggplot(df) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)

ggplot(df) + geom_line(aes(index, log(value))) + facet_wrap(~ variable, scale="free", ncol=1)
```

d.p <- c(5, 5)
s2.p <- c(5, 5)

s2_prior <- function(x) dgamma(x, s2.p[1], s2.p[2])
d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2])
prior_fns <- list(l = d_prior, sigma.n = s2_prior)

require(plyr)
prior_curves <- ddply(df, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100)
    data.frame(value = grid, density = prior_fns[[dd$variable[1]]](grid))
})

# Posteriors (easier to read than histograms)
ggplot(df, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

```{r}
prior_curves <- ddply(df, "variable", function(dd){
    grid <- seq(-4, 4, length = 100)
    data.frame(value = grid, density = prior_fns[[dd$variable[1]]](grid))
})

ggplot(prior_curves, aes(value, density)) + 
  geom_line(data=prior_curves, col="red") + 
  facet_wrap(~ variable, ncol=1, scale="free_y")
```
